---
title: "oRaklE: Multi-Horizon Electricity Demand Forecasting in High Resolution"
author: "Johannes Schwenzer, Simone Maxand, Tatiana C. G. Grandón"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{oRaklE: Multi-Horizon Electricity Demand Forecasting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The **oRaklE** R package provides tools for long-term electricity demand forecasting with hourly resolution on a country-wide scale. Designed for scalability and precision, it supports stepwise modeling of electricity demand components (long-term, mid-term, and short-term) and enables detailed analysis and forecasting. 

This vignette provides a comprehensive overview of the package's features and demonstrates how to use its functionalities. For better displaying quality we recommend visiting 

---

# Package Workflow

The package's functions can be used independently or combined using the `full_forecast()` function. Below is the flowchart that outlines the package workflow:

```{r, echo=FALSE, out.width='400px', fig.align='center'}
knitr::include_graphics("https://github.com/user-attachments/assets/26ffb2f7-de0c-471e-8226-f133a095ee41")
```


---

# Usage

## Step 1: Data Acquisition and Preparation

As a first step, the package gets the load data from the Transparency Platform of the European Network of Transmission System Operators for Electricity (ENTSO-E) https://transparency.entsoe.eu/. If the country is not a member of the ENTSO-E a dataset needs to be supplied manually. 

### Retrieve Load Data:
```r
demand_data <- get_entsoE_data(2017, 2021, "France")
```
```{r, echo=FALSE, out.width='400px', fig.align='center'}
knitr::include_graphics("https://github.com/user-attachments/assets/24450818-c869-4142-a5b3-cb2a2ec7ff36")
```

### Fill Missing Data:
The next step checks for missing data and replaces NaN values with the load data one week prior at the same time. So a missing value on a Thursday at 8 pm will be filled with the load value at 8 pm the week before. It also adjusts the dataset to account for changes in time resolution, ensuring that the data is standardized to an hourly resolution. For example, French load data is reported at an hourly resolution from 2017 to 2021 and at a half-hourly resolution from 2022 onwards. Therefore, it is recommended to run this function, even if there are no missing values.

```r
demand_data_filled <- fill_missing_data(demand_data)
```

---

## Step 2: Decomposition of Load Data
After the data is downloaded and standardized, the load time series is decomposed into three components: a yearly long-term trend, a daily mid-term seasonality, and an hourly short-term seasonality. If the data is available only at a daily resolution, the calculation of hourly seasonality is skipped.



```r
decomposed_data <- decompose_load_data(demand_data_filled)
```
```{r, echo=FALSE, out.width='400px', fig.align='center'}
knitr::include_graphics("https://github.com/user-attachments/assets/e5db1014-6e2b-4632-ab00-5923e8414553")
```

The function returns a list of three dataframes —one for each time series component:
 - **demand_data$longterm**
 - **demand_data$midterm**
 - **demand_data$shortterm**
 
In the following steps, each time series will be modelled individually. 
---

## Step 3: Long-Term Trend Modeling

### Retrieve Historical Data:
First, historical data from the ENTSO-E archive, starting from 2006, is added to the long-term trend series. A dataset with historic yearly demand data for each ENTSOE-E member country is included in the library.
```r
longterm <- get_historic_load_data(decomposed_data$longterm)
```

### Add Macroeconomic Covariates:
The long-term electricity demand trend is estimated with different regression algorithms based on macro-economic covariates. 10 macro-economic indicators are fetched via an API call to the Worldbank Development Indicators. Additional macro-economic covariates can be added manually.
```r
longterm_all_data <- get_macro_economic_data(longterm)

head(longterm_all_data)
#   country   year  avg_hourly_demand    population        GDP         industrial_value_added   ...
#   FR        2006        54417.12        63628261     2.279283e+12           19.28408
#   FR        2007        54769.12        64021737     2.334550e+12           19.13674
#   FR        2008        56214.75        64379696     2.340502e+12           18.81383
#   FR        2009        55409.97        64710879     2.273252e+12           18.30484
#   ...
```
It should be noted that the average hourly demand (avg_hourly_demand) refers to the average demand over each hour of the respective year (typically in MW). To calculate the total annual demand, multiply avg_hourly_demand by 8760 hours.


### Derive Long-Term Prediction Models:

After the dataset is fully prepared the best long-term prediction models are derived with multiple linear regression and k-fold cross-validation. Details on the mathematical approach are specified in the accompanying paper. The variable for test_set_steps defines how many years are used for the test set (also commonly referred to as validation set). The testquant variable defines how many of the initial best models are subjected to cross-validation.
```r
longterm_predictions <- long_term_lm(longterm_all_data, test_set_steps = 2, testquant = 500)
```
The three best models as well as plots for each model are generated and saved.

```{r, echo=FALSE, out.width='400px', fig.align='center'}
knitr::include_graphics("https://github.com/user-attachments/assets/3facab6e-5e7c-4f6a-8e13-c53d6d9591a0")
```
### Generate Future Predictions:
Once the best models are derived, future predictions can be made. Since these models rely on macroeconomic covariates, which are unknown for future years, forecasts for these indicators are needed. The library can obtain these forecasts from the [World Economic Outlook Database](https://www.imf.org/en/Publications/WEO/weo-database/2023/October) of the International Monetary Fund, or users can manually include predictions from other sources or specific scenarios.

The *end_year* variable specifies until which year the predictions will be made. If the *WEO* dataset is used, predictions can only be made up to 2028. If the dataset variable is set to any option other than WEO, the function will prepare the data frame structure and list the required macroeconomic indicators. 
```r
## Prepare dataset for future predictions
# With the dataset option set to the default ("WEO")
longterm_future_macro_data <- long_term_future_data(longterm_predictions, end_year = 2028, dataset = "WEO")
tail(longterm_future_macro_data)
#   country   year   avg_hourly_demand     population          GDP           industrial_value_added   ...
#     FR      2021          53225.29        67764304      2.575192e+12               16.39563
#     FR      2022             NA           71764951      2.640147e+12               17.56864
#     FR      2023             NA           75808393      2.665256e+12               17.73190
#     FR      2024             NA           77673096      2.701124e+12               17.71412
#     FR      2025             NA           79192297      2.750043e+12               18.21028
#     FR      2026             NA           80760586      2.795947e+12               18.48387
#     FR      2027             NA           82230070      2.838581e+12               18.72595
#     FR      2028             NA           83540334      2.879402e+12               19.03256

# With the dataset option set to anything else (e.g., "manual")
longterm_future_macro_data <- long_term_future_data(longterm_predictions, end_year = 2028, dataset = "manual")

# Output:
# If you want to use your own dataset you will need predictions for the following macro-economic variables:
# 
#  GDP, GNI, industrial_value_added, rural_population for model 1 
# 
#  GDP_growth, household_consumption_expenditure, rural_population, service_value_added for model 2 
# 
#  GDP, industrial_value_added, rural_population, service_value_added for model 3 

```
After the dataset for the future predictions is prepared, long-term forecasts until the designated time period can be made. One forecast with each of the three best models is calculated and the results are saved and plotted.
```r
longterm_future_predictions <- long_term_future(longterm_future_macro_data)
```
```{r, echo=FALSE, out.width='400px', fig.align='center'}
knitr::include_graphics("https://github.com/user-attachments/assets/6ea76ae8-5aea-4d3b-9c48-132da3ebc269")
```
---

## Step 4: Mid-Term Seasonality Modeling

## Mid-Term Component

The mid-term component is the difference between the yearly hourly average demand and the daily average hourly demand of the respective day:

$$
D_M(y,d) = D_m(d) - D_L(y)
$$

where:

- \(D_M(y,d)\) refers to the mid-term component at day \(d\) and year \(y\),
- \(D_m(d)\) refers to the average daily hourly load of day \(d\),
- \(D_L(y)\) refers to the yearly average hourly load of year \(y\).

The mid-term time series is modeled using seasonal, calendar, and temperature-related variables. Seasonal covariates include:

- **Month** (January–December),
- **Day of the week** (Sunday–Saturday),
- **Holiday indicator**: A dummy variable indicating whether the day is a holiday or a workday.

Information about public holidays is retrieved from [https://date.nager.at](https://date.nager.at).


### Add Holidays:
```r
midterm_demand_data <- add_holidays_mid_term(decomposed_data$midterm)
```

### Retrieve Weather Data:
```r
midterm_demand_and_weather_data <- get_weather_data(midterm_demand_data)
```

### Derive Mid-Term Prediction Models:
```r
midterm_predictions <- mid_term_lm(midterm_demand_and_weather_data$demand, 
                                    test_set_steps = 730, Tref = 18, 
                                    method = "temperature transformation")
```

### Generate Future Predictions:
```r
midterm_future_predictions <- mid_term_future(midterm_predictions, end_year = 2028)
```

---

## Step 5: Short-Term Seasonality Modeling

### Add Holidays:
```r
shortterm_demand_data <- add_holidays_short_term(decomposed_data$shortterm)
```

### Derive Short-Term Prediction Models:
```r
shortterm_predictions <- short_term_lm(shortterm_demand_data, test_set_steps = 17520)
```

### Generate Future Predictions:
```r
shortterm_future_predictions <- short_term_future(shortterm_predictions, end_year = 2028)
```

---

## Step 6: Combine All Models

Finally, the individual models are combined into a complete demand forecast:

```r
full_model_predictions <- combine_models(longterm_predictions, 
                                         midterm_predictions, 
                                         shortterm_predictions, 
                                         longterm_model_number = 1)
```

### Generate Future Forecasts:
```r
full_model_future_predictions <- combine_models_future(longterm_future_predictions, 
                                                       midterm_future_predictions, 
                                                       shortterm_future_predictions, 
                                                       longterm_model_number = 1)
```

---

# All-in-One Function

The `full_forecast()` function automates all the steps described above. Example usage:

```r
forecast_data <- full_forecast(start_year = 2017, end_year = 2021, country = "France", 
                               test_set_steps = 2, future = "yes", end_year = 2028)
```


